---
title: "Jail Population Growth and Equality Implications"
author: "Ryan Breuer"
date: '2022-11-01'
output: html_document
---


```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/analysis.r")
source("../source/a4-helpers.R")
```

## Abstract

This report addresses discrepancies across a variety of countries related to jailed populations. There are unjust power structures that exist in our society and data science has the power to reveal these systems. This data impacts populations across a wide range of races, communities, and groups. Communities in which the jailed populations are located and their families are some of the indirect stakeholders of this set. The questions that I answer in my report include how people of different races are represented in the jailed populations of our country. 


## Section 1: Introduction

This report addresses social issues of racial inequality, biases, and discrepancies across the countries related to jailed populations. There are power structures that exist and in order to combat them, we have to use the data and resources we have access to in order to reveal unjust systems. The direct stakeholders of this data set are those in the populations included in this data set. The indirect stakeholders impacted by this data set include the communities in which the jailed populations are located and their families. The questions that I answer in my report include how people of different races are represented in the jailed populations of our country.

## Section 2: Data Summary 

I decided to calculate data regarding information surrounding the black juvenile jail population in my home county during my birth year. I saw that there was a ~20% drop in this population in the fist year, a ~15% rise in the second year, and a ~20% rise in the third year after my birth. I found this information very intriguing and decided to dive deeper into it. I discovered that there are patterns of inequality in Maricopa County and am interested to explore this data further. My hypothesis is that the black population was impacted by policies in which targeted juveniles.

# Section 3: Jail Population Rise in the United States

The chart below shows the exponential increase in the jail population in the United States over the past 30 years. This data is very revealing in how more and more people are being jailed every year. This may have implications on how policies are enacted and how populations are committing more crime. 


```{r echo=FALSE}
plot(plot_jail_pop_for_us)
```

# Section 4: Jail Population Rise by States

This chart below showcases the varying rise in jail populations across the country. It is interesting that there are large discrepancies in how the population has experienced growth in different states. This graph is foccussed on the past 30 years.


```{r echo=FALSE}
plot(plot_jail_pop_by_states)
```

## Summary (Section 4):

This data shows us that there are systematic changes in which are increasing the amount of people going to jail. Although the population is rising in the United States, the jailed population is rising at a higher rate. This data tells us that we need to investigate our justice system and understand why this population is rising. 

# Section 5: Black vs. White Jail Population Growth in Washington Overtime

This chart below shows the difference in black and white population growth in Washington jails. It shows that there has been a recent increase in the population of black people vs. white. The main takeaway is that there is a trend for more black people (in ratio to white) over time.


```{r echo=FALSE}
plot(balck_vs_black_jail_pop_per_state)
```

## Summary (Section 5):
This data shows that there has been an exponential increase in the population growth of black people in Jail in Washington. The alpha is the portion of population growth that is equalized by the two races.

# Section 6: Map Showcasing Percantage of Each State's Black Jail Population

This map displays how there are varying percentages of state's black population in jails. As you can see, there is a higher proportion in the south east vs the northwest. My hypothesis is that this is due to racial bias from historical slavery in these areas.  


```{r echo=FALSE}
plot(map)
```


